\input{../style/cs421_style.tex}
\usepackage{algorithm}
\usepackage{listings}
%\usepackage{algpseudocode}
\usepackage{graphicx,amssymb,amsmath}
\usepackage{epstopdf}
\sloppy

\oddsidemargin -0.5in
\evensidemargin -0.5in
\textwidth 7.5in
\topmargin -0.5in
\textheight 9.0in

\begin{document}

\solution{Dan McQuillan}{\today}{6}{Summer 2015}

\pagestyle{myheadings}  % Leave this command alone
	 	
\section{Theoretical Questions}

\subsection{Problem 1}

In both cases: \\
let \( \Sigma = \{ a, b, c \} \)

\subsubsection{The language where a occurs in every third position}

\def \booleanOr { \( \vee \) }
\def \abcDisjunction { (a \( \vee \) b \( \vee \) c) }
( \abcDisjunction \abcDisjunction a)* ( ( \abcDisjunction \abcDisjunction a ) \booleanOr ( \abcDisjunction \abcDisjunction ) \booleanOr \abcDisjunction )? 

\subsubsection{The language where each string contains exactly 3 c's}
(a \(\vee\) b)* c (a \(\vee\) b)* c (a \(\vee\) b)* c (a \(\vee\) b)*

\subsection{Problem 2}

\begin{enumerate}
	\item\(L_1 = \{ w | w \text{ starts with a symbol 0 and contains the symbol 1 at least once} \} \text{ where } \Sigma = \{ 0, 1 \} \)
	% regular
	
	\begin{enumerate}
		\item{Regular expression}
				
		0 ( 0 \(\vee\) 1 )* 1 ( 0 \(\vee\) 1 )*		
				
		\item{Regular grammar}
		
		\( <Language\_1> :: = 0 <ZeroOrOne> \) \\
		\( <ZeroOrOne>   :: = 0 <ZeroOrOne> \) \\
		\( <ZeroOrOne>   :: = 1 <ZeroOrOneOrEmpty> \) \\
		\( <ZeroOrOneOrEmpty> :: = 0 <ZeroOrOneOrEmpty> \) \\
		\( <ZeroOrOneOrEmpty> :: = 1 <ZeroOrOneOrEmpty> \) \\
		\( <ZeroOrOneOrEmpty> :: = \epsilon \)
	\end{enumerate}
	
	\item\(L_2 = \{ w | w \text{ contains an equal number of 0s and 1s} \} \text{ where } \Sigma = \{ 0, 1 \} \)
	% non-regular
	
	\begin{proof}
	By contradiction; assume the language \(L_2\) is regular. 
	
	Let n be the length guaranteed by the pumping lemma. Suppose we have a string \( w = 0^n 1^n \).  Then \( | w | = 2 n \geq n \) and \( w \in L_2 \).  
	
	Therefore, there exists strings x, y, and z such that w = xyz, \( | xy | \leq n, y \neq \epsilon \) and for any number i, \( xy^iz \in L_2 \). 
	
	Since, \( | x y | \leq n \) y must consist of only 0s.  However \(xy^2z = 0^{n + |y|}1^n \), and since \( |y| > 0 \), we have that \(xy^2z \notin L_2 \)
	
	Therefore, we have a contradiction and our language is not regular.
	\end{proof}
	
	\item\(L_3 = \{ w | \text{ the length of w is odd} \} \text{ where } \Sigma = \{ a, b \} \)
	% regular
	
	\begin{enumerate}
		\item{Regular expression}
		
		( a \(\vee\) b ) ( ( a \(\vee\)  b ) ( a \(\vee\) b ) )*
		\item{Regular grammar}
		
		\( <Language\_3> :: = a <AOrBOrEmpty> \)\\
		\( <Language\_3> :: = b <AOrBOrEmpty> \)\\
		\( <AOrBOrEmpty> :: = b <Language\_3> \)\\
		\( <AOrBOrEmpty> :: = a <AOrC> \)\\
		\( <AOrBOrEmpty> :: = \epsilon \)
	\end{enumerate}
	
	\item\(L_4 = \{ w | w \text{ does not contain symbol a immediately followed by symbol b} \} \text{ where } \Sigma = \{ a, b, c \} \)
		
	% non-regular
	
	\begin{enumerate}
		\item{Regular expression}
		
		((aa)*(ac) \(\vee\) b \(\vee\) c)* a
		\item{Regular grammar}
		
		\( <Language\_4> ::= a <AOrC> \)\\
		\( <Language\_4> ::= b <Language\_4> \)\\
		\( <Language\_4> ::= c <Language\_4> \)\\
		\( <Language\_4> ::= \epsilon \)\\
		\( <AOrC> ::= c <Language\_4> \)\\
		\( <AOrC> ::= a <AOrC> \)\\
		\( <AOrC> ::= \epsilon \)\\
		
	\end{enumerate}
	
	\item\(L_5 = \{ w | \text{ the length of w is a perfect cube } \} \text{ where } \Sigma = \{ a, b, c \} \)
	% non-regular
	
	\begin{proof}
	By contradiction; assume the language \(L_5\) is regular.
	
	Let n be the length guaranteed by the pumping lemma.   
	
	Suppose \( w = a^{\frac{n^3}{3}} b^{\frac{n^3}{3}} c^{\frac{n^3}{3}} \).
	
	Since \( |w| = \frac{n^3}{3} + \frac{n^3}{3} + \frac{n^3}{3} = n^3 \), which is a perfect cube, \( w \in L_5 \).
	
	By the pumping lemma we know that we can split \( w = xyz \) s.t. the conditions of the pumping lemma hold.
	
	We know that:
	\[
		1 \leq | y | \leq | xy | \leq n
	\]
	
	Since for the pumping lemma to hold we also require that any amount of y terms in the middle to still let the expression hold. Therefore we know also that:
	\[
		xy^2z \in L_5
	\]
	
	Therefore we may assume that \( | xy^2z | \) is a perfect cube.  However we know that:
	
	\begin{align}
		n^3 {}& = | w | \\
		       {}& = | xyz | \\
		       {}& < | xy^2z | \\
		       {}& \leq n^3 + n \hspace*{8mm}\text{ since, } | y | \leq n \\
		       {}& < n^3 + 3n^2 + 3n + 1
	\end{align}
	
	In summary, we now know that:
	\[
		n^3 < | xy^2z | < n^3 + 3n^2 + 3n + 1
	\]
	
	That is \( | xy^2z | \) lies between two subsequent perfect cubes.  Therefore, it cannot be a perfect cube itself, and hence we have a contradiction to \( xyyz \in L_5 \).
	\end{proof}
\end{enumerate}

\end{document}

